Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-30 paper/paper.tex
replace abstracting with abstract

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Abstracting abstract machines (AAM) is a systematic methodology for constructing 
  abstract interpreters that are derived from concrete small-step abstract machines.
  \citeauthor{darais2017abstracting} apply the same idea on monadic definitional interpreters, 
  and obtain monadic abstract definitional interpreters (ADI) in big-step style.
  Yet, the relations between small-step abstracting abstract machines and big-step 
  abstracting definitional interpreters are not well studied.

@@

\subsection{Abstracting Abstract Machines} \label{aam}
Abstracting abstract machines (AAM) is a systematic methodology that derives sound
abstract interpreters for higher-order functional languages from concrete abstract 
machines \cite{van2012systematic, van2010abstracting}. An abstracting abstract machine
implements computable abstract semantics which approximates the runtime behaviors of 
programs.
Since the state space of concrete execution is possibly infinite, the key insight of 
AAM approach when analyzing programs is to allocate both bindings and continuations 
on the store, and then bound the addresses space to be finite. Since each component
of state is finite, the abstracted machine-state space is also finite, and therefore
computable.

In this section, we derive the abstracting abstract machine from concrete
CESK machines, and also show how to instantiate useful $k$-call-sensitive 
control-flow analysis.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstracting
    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
    describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates \cite{friedman-ai}.
    In this case, nondeterminism can be easily handled via @for@ comprehension. 
\end{itemize}

@@

\textit{Pushdown Control-Flow Analysis.}
In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return match based on small-step
abstracting abstract machines.

@@

\textit{Abstracting Definitional Interpreters.}
Reynold's seminal paper \emph{Definitional interpreters for higher-order programming languages} 
\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in some definitional interpreters, the
defined language can inherit properties from the defining language. With this insight,
\citeauthor{darais2017abstracting} construct abstracting definitional interpreters which
automatically inherit the pushdown control-flow property from its defining language because
the defined language simply uses the call-stack model of the meta-language~\cite{darais2017abstracting}.
\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work on direct-style
$\lambda$-calculus, and itself is written in monadic style. One of the advantages of monadic
abstract interpreters is modular and composable, therefore deploying different sensitivities
or features is just applying a different monad. Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979}
show a monadic abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Abstracting abstract machines (AAM) is a systematic methodology for constructing 
  abstract interpreters that are derived from concrete small-step abstract machines.
  \citeauthor{darais2017abstracting} apply the same idea on monadic definitional interpreters, 
  and obtain monadic abstract definitional interpreters (ADI) in big-step style.
  Yet, the relations between small-step abstract abstract machines and big-step 
  abstract definitional interpreters are not well studied.

@@

\subsection{Abstracting Abstract Machines} \label{aam}
Abstracting abstract machines (AAM) is a systematic methodology that derives sound
abstract interpreters for higher-order functional languages from concrete abstract 
machines \cite{van2012systematic, van2010abstracting}. An abstract abstract machine
implements computable abstract semantics which approximates the runtime behaviors of 
programs.
Since the state space of concrete execution is possibly infinite, the key insight of 
AAM approach when analyzing programs is to allocate both bindings and continuations 
on the store, and then bound the addresses space to be finite. Since each component
of state is finite, the abstracted machine-state space is also finite, and therefore
computable.

In this section, we derive the abstract abstract machine from concrete
CESK machines, and also show how to instantiate useful $k$-call-sensitive 
control-flow analysis.

@@

We have multiple choices of how to proceed:
\begin{itemize}
  \item By monadifying the continuations, we obtain the abstract definitional interpreters 
    in monadic style which are what \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates \cite{friedman-ai}.
    In this case, nondeterminism can be easily handled via @for@ comprehension. 
\end{itemize}

@@

\textit{Pushdown Control-Flow Analysis.}
In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
gilray2016pushdown, johnson2015abstracting} to achieve precise call/return match based on small-step
abstract abstract machines.

@@

\textit{Abstracting Definitional Interpreters.}
Reynold's seminal paper \emph{Definitional interpreters for higher-order programming languages} 
\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in some definitional interpreters, the
defined language can inherit properties from the defining language. With this insight,
\citeauthor{darais2017abstracting} construct abstract definitional interpreters which
automatically inherit the pushdown control-flow property from its defining language because
the defined language simply uses the call-stack model of the meta-language~\cite{darais2017abstracting}.
\citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style
$\lambda$-calculus, and itself is written in monadic style. One of the advantages of monadic
abstract interpreters is modular and composable, therefore deploying different sensitivities
or features is just applying a different monad. Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979}
show a monadic abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.
